Experimentally informed site-specific substitution models deepen phylogenetic estimates of the divergence of viral lineages

### 2018-02-14 manuscript/manuscript.tex
changes to simulations section

Next, we wanted to test the effect of substitution model on branch length estimation given sequences with site-specific amino-acid frequencies. 
To this end, we simulated sequences under an ExpCM and re-inferred the branch lengths using the models  in \ref{fig:model_feature}C.

Models with uniform stationary state underestimated the branch lengths of the simulated sequences. 
This underestimation is most pronounced on the long branches, with the GY94 estimating branches which are only $\sim \frac{2}{3}$ of the true branch length (\ref{fig:simulations}).
Accounting for the site-to-site rate variation by $\Gamma\omega$ does not prevent this underestimation as the GY94+$\Gamma\omega$ estimates roughly the same branch length as the GY94. 
We also estimated branch lengths using an average ExpCM. 
These models have a uniform stationary state defined by the average preference of amino acid across all sites. 
Even these models, which share some information with the model used to simulate the sequences, underestimate the long branches.

As expected neither the ExpCM nor the ExpCM+$\Gamma\omega$ underestimate the long branches. 
However, the variance between simulations increases as the branch length increases, though this error is not biased towards over- or underestimation. 
This increase in error despite a perfect model match simply underscores the difficulty of estimating extremely long branches. 
Despite this universal difficulty, it is clear that models with uniform stationary states will always drastically underestimate the long branch lengths.

>>>>

Next, we tested the effect of stationary state and $\Gamma\omega$ rate variation on branch length estimation given sequences with site-specific amino-acid frequencies. 
We simulated sequences under an ExpCM, re-estimated the branch lengths using the models in \ref{fig:model_feature}C, and compared these estimates to the actual branch lengths.

The models with a uniform stationary state underestimated the branch lengths of the simulated sequences, especially the lengths of long branches. 
The GY94 model estimated branches lengths which are only about 60\% of the true branch length for the longest branches. 
Accounting for site-to-site rate variation by $\Gamma\omega$ did not prevent the underestimation of long branches, as the GY94+$\Gamma\omega$ inferred approximately the same branch lengths as the GY94 model.
We also estimated branch lengths using averaged ExpCMs.
The averaged ExpCMs have a uniform stationary state defined by the average DMS preference for a given amino acid across all sites from the deep mutational scan. 
Even these models, which share information with the model used to simulate the sequences, underestimated the long branches. 
The inability of these uniform stationary state models to specifically describe purifying selection on a site by site basis clearly hinder their ability to accurately estimate long branch lengths.

Neither the ExpCM nor the ExpCM+$\Gamma\omega$ underestimated the long branches. 
This is an unsurprising result, as the sequences were simulated under an ExpCM. 
However, while not biased towards over- or underestimation, the variance between simulations increased as the branch length increased. 
The increase in error, even with a perfect model match, underscores the difficulty of estimating extremely long branches. 
Despite this universal difficulty, it is clear from the GY94 and averaged ExpCM results that models with a uniform stationary state will always drastically underestimate long branches.
